---
output:
  html_document: default
  pdf_document: default
---
Red Wine by Ute Stohner
========================================================

load.package(ggthemes)

```{r echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE)
# Load all of the packages that you end up using in your analysis in this code
# chunk.
library(ggplot2)
library(dplyr)
library(GGally)
library(scales)
library(memisc)
library(tidyverse)
library(modelr)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
RedWine <- read.csv('wineQualityReds.csv', header=TRUE, sep=",")
```

#Univariate Plots

Getting overview first

```{r echo=FALSE}

str(RedWine)

```



```{r echo=FALSE}

apply(RedWine, 2, function(x) any(is.na(x)))

```


Creating summaries for other variables

```{r echo=FALSE}

summary(RedWine$X)
summary(RedWine$fixed.acidity)
summary(RedWine$volatile..acidity)
summary(RedWine$citric.acid)

```

```{r echo=FALSE}

summary(RedWine$sulphates)
summary(RedWine$residual.sugar)
summary(RedWine$chlorides)
summary(RedWine$free.sulfur.dioxide)
summary(RedWine$total.sulfur.dioxide)

```


```{r echo=FALSE}

summary(RedWine$sulphates)
summary(RedWine$alcohol)
summary(RedWine$density)
summary(RedWine$pH)

```

```{r echo=FALSE}

summary(RedWine$alcohol)
summary(RedWine$quality)

```


Getting familiar with the file.

```{r echo=FALSE}
names(RedWine)
head(RedWine)
```

I am assuming that quality is considered to be the  dependent variable that
we are trying to influence.
Goal is to understand what influences wine quality.
Not knowing much about  wine making I could imagine that this helps 
wine producers to understand where to grow wine, which grapes to grow 
and also how to blend wine. 
Qualitiy is a qualititave (ordinal) value. 
All other values are quantitative.

## Examining quality 

What is the scale for quality?

```{r}
unique(RedWine$quality)
```

Finding out where most wines fall



```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(aes(x=quality), data=RedWine)+geom_histogram(binwidth=1, color='white')+
scale_x_continuous(breaks=seq(0,10,1))

```

Most wines have a quality of 5 and 6

Alcohol varies from 8.4 to 14.9. Median at 10.2

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Histogram for alcohol
ggplot(aes(x=alcohol), data=RedWine)+geom_histogram(binwidth=0.3)

```

Does sugar influence alcohol content?

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Grouping by alcohol and getting mean sugar per alcohol
by_alcohol2<-group_by(RedWine,alcohol)
RW_byalcohol2<-summarise(by_alcohol2, mean_sugar=mean(residual.sugar))

```

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Adding the results from above into a line plot
ggplot(aes(x=alcohol, y=mean_sugar), data=RW_byalcohol2)+geom_line()+xlim(8,14)

```


Does the graph look better with rounded alcohol?



Spike just at over 8, probably caused by an outlier in high sugar.
How does it look with rounded numbers?

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Rounding alcohol
RedWine$alcohol_round<-round(RedWine$alcohol)

#Sane grouping, but with rounded alcohol
by_alcohol<-group_by(RedWine,alcohol_round)
RW_byalcohol<-summarise(by_alcohol, mean_sugar=mean(residual.sugar))

```

```{r echo=FALSE,warning=FALSE,message=FALSE}


ggplot(aes(x=alcohol_round, y=mean_sugar), data=RW_byalcohol)+geom_line()+
xlim(8,14)

```

Alcohol does not rise or fall with higher sugar levels
Creating alcohol_level to summarise alcohol for further plots.

```{r echo=FALSE,warning=FALSE,message=FALSE}
#creating alcohol level
RedWine$alcohol_level<-cut(RedWine$alcohol_round,c(8.4,10,12,14.9),
labels = c("Low (8.4 to 10)", "Medium (10 to 12)",
"High (12 to 14.9)"))

```

Volatile acidity and Quality

```{r echo=FALSE,warning=FALSE,message=FALSE}

ggplot(aes(x=quality,y=volatile.acidity), data=RedWine)+geom_jitter(alpha=1/10)+
geom_line(stat='summary', fun.y=mean)

```


With higher quality, volatile acidity goes down.

Fixed.acidity and volatile.acidity


According to this article:
http://winemakersacademy.com/understanding-wine-acidity/
most wine drinkers perceive the total amount of acidity
and not the volatile and fixed acidity separate. Adding a new column for 
total acidity

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Creating total acidity
RedWine$total.acidity=RedWine$volatile.acidity+RedWine$fixed.acidity


```


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Histogram for total acidity
ggplot(aes(x=total.acidity), data=RedWine)+geom_histogram(binwidth=0.3)


```

How does acidity influence quality?

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Scatterplot for total acidity and quality
ggplot(aes(y=total.acidity, x=quality), data=RedWine)+geom_jitter(alpha=1/10, position=position_jitter(h=0))

```

Removing outliers in high acidity

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Removing top 3 % from the scatterplot above and adding line for total acidity 
#mean
ggplot(aes(y=total.acidity, x=quality), data=subset(RedWine, total.acidity<=quantile(RedWine$total.acidity, 0.97)))+geom_jitter(alpha=1/10, position=position_jitter(h=0))+
geom_line(stat='summary', fun.y=mean)

```
Not a clear picture for total acidiy and quality

Same plot with volatile acidity plus line for the mean

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Same plot with volatile instaed of total acidity
ggplot(aes(y=volatile.acidity, x=quality), data=subset(RedWine, volatile.acidity<=quantile(RedWine$volatile.acidity, 0.97)))+
geom_jitter(alpha=1/10, position=position_jitter(h=0))+
geom_line(stat='summary', fun.y=mean)

```

Low volatile.acidity influences quality.

Adding in quality_factor, to have the data as a factor


```{r echo=FALSE,warning=FALSE,message=FALSE}

RedWine$quality_factor<-factor(RedWine$quality)

```


Frequency polygon for acidity colored by quality factor



```{r echo=FALSE,warning=FALSE,message=FALSE}

#Frequency polygon for volatile acidity, using quality_factor
#for separating qualities
ggplot(aes(x=volatile.acidity), data=RedWine)+
geom_freqpoly(aes(color=quality_factor))+
scale_color_brewer(type = "div",
palette = "RdYlBu", name="Quality", direction=-1)+
theme_dark()

```

This graph is not very helpful, mainly as the numbers in our 
quality distribution are so different.
Most wines fall into qualtiy 5 and 6. Trying a log scale

Frequency polygon for acidty colored by quality, log scale

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Adding in log scale
ggplot(aes(x=volatile.acidity), data=RedWine)+
geom_freqpoly(aes(color=quality_factor))+scale_y_log10()+
scale_color_brewer(type = "div",
palette = "RdYlBu", name="Quality", direction=-1)+
theme_dark()

```

With the log scale, the problem of the small size of high and low quality still 
persists. 


pH and acidity are connected. http://chemistry.elmhurst.edu/vchembook/184ph.html
Acidity starts at ph level 7. In the RedWine data ph ranges from 2.7 to 4.0.
Selecting total.acidity and ph

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Scatterplot for pH and total acidity
ggplot(aes(y=pH, x=total.acidity),data=RedWine)+geom_jitter(alpha=1/10)+
scale_y_continuous(breaks=c(3,4))+stat_smooth(method ='lm')

```


Checking volatile and fixed acidity separately.
Starting with volatile acidity.


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for volatile acidity
ggplot(aes(y=pH, x=volatile.acidity),data=RedWine)+
geom_jitter(alpha=1/10)+
scale_y_continuous(breaks=c(3,4))+stat_smooth(method ='lm')

```


Not the same picture as with total acidity. 
pH is going up  with volatile acidity.

Fixed acidity

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for fixed acidity
ggplot(aes(y=pH, x=fixed.acidity),data=RedWine)+geom_jitter(alpha=1/10)+
scale_y_continuous(breaks=c(3,4))+stat_smooth(method ='lm')

```

pH is going down with fixed acidity.
It is not a good idea to combine acidity, as fixed and volatile acidity
seem to be different.


Quality and pH

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Quality and ph Scatterplot
ggplot(aes(y=pH, x=quality),data=RedWine)+geom_jitter(alpha=1/5)+
scale_y_continuous(breaks=c(3,4))

```

The graph does not show a strong correlation between pH and quality.


Density and Alcohol

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for Density and alcohol, adding in summary line for
#median density
ggplot(aes(x=quality,y=density),data=RedWine)+geom_jitter(alpha=1/5, position=position_jitter(h=0))+scale_y_continuous(limits=c(0.9925, 1))+
stat_summary(geom = 'line', fun.y = median)

```

Lower Density means higher quality.But it does not look like a 
strong influence.
Again we have the problem of the uneven distribution 
of wines in the mediuam (5/6) quality section.


Distribution of Sulphates

```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(aes(x=sulphates), data=RedWine)+geom_histogram(binwidth = 0.1)+
scale_x_continuous(breaks=seq(0,2,0.2))
```

Most wines have sulphates of around 0.6, sulphates fall rapidly, but
some outliers of high sulphate wines.

Distribution of citric acid

```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(aes(x=citric.acid), data=RedWine)+geom_histogram(binwidht=0.3)
```

Most citric acid values fall between 0 and 0.420 (quartile spread).
The distribution is not even and shows outliers.



### What is the structure of your dataset?
1599 obs. of  14 variables.
Quality is ordinal with valus from 3 to 8. 8 being the best quality.

Most wines fall into quality levels of 5 and 6
Median alcohol is 10.2

### What is/are the main feature(s) of interest in your dataset?
The main feature is quality.
So far I can not tell what determines wine quality. 
Assumption is, it is a combination of variables.

### What other features in the dataset do you think will help support your \
It is too early to say. At this point I would concentrate on a combination of 
alcohol, acidity, ph and density. Need more investigation

### Did you create any new variables from existing variables in the dataset?
Total acidity was created. alcohol_round and quality_factor were created.


### Of the features you investigated, were there any unusual distributions? 
Looking at the number of wines, the majority of wines falls in 5 and 6.
I would have assumed that volatile and total acidity can be combined.
But comparing with pH, they seem to have different influence on the wine.

# Bivariate plots



Running ggpairs

```{r  fig.width=12, fig.height=9, Bivariate_Plots_ggpairs }

ggpairs(RedWine,
  lower = list(continuous = wrap("points",size = 1)),
  upper = list(combo = wrap("box", size = 1))) +
  theme_minimal(8)

```



##Correlation for quality

Focusing first on the correlation for quality. Only two variables have 
a weak influence on quality
Quality/Volatile acidity: -0.391
Quality/Acohol: 0.476


```{r echo=FALSE,warning=FALSE,message=FALSE}

cor(RedWine$quality, y=RedWine$volatile.acidity, method = c("pearson"))
cor(RedWine$quality, y=RedWine$alcohol, method = c("pearson"))

```

## Volatile acidity and quality

After exploring single variables in the last section, I would like to understand
if there is a combination of variables that influences wine quality.
The correlation coefficient suggests there is a weak influence of both 
volatile acidity and alcohol.

Boxplot for volatile.acidity and quality

```{r echo=FALSE,warning=FALSE,message=FALSE}


#Boxplot for volatile acidity and quality
ggplot(aes(y=volatile.acidity, x=quality_factor),data=RedWine)+
geom_boxplot()

```

The box plot shows that the acidity falls with higher quality.
The spread of of data gets smaller as the quality increases. 
We find the smallest spread  for quality 8, the highest for quality 4. 
It also shows that 5 and 6 have more outliers.
It will be a good idea to remove outliers for volatile acidity in the next 
graphs.


Combining alcohol, quality and volatile acidity in one plot.



```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for volatile acidity and quality. Points colored by alcohol level
ggplot(aes(y=volatile.acidity, x=quality), 
#Removing na acohol data
data=subset(RedWine, !is.na(alcohol_level) ))+
geom_jitter(alpha=1/2,position=position_jitter(h=0),
aes(color=alcohol_level))+
geom_line(stat='summary', fun.y=mean)+ylim(0,1.2)+
#Changing color palette
scale_colour_brewer(palette="Set1", "Alcohol Level", direction=-1)
       
```


While this plot shows that the acidity falls in higher quality while  alcohol 
rises, it does not show which combination of alcohol and acidity is 
ideal.

The correlation coefficient of Alcohol and volatile acidity does not suggest 
a connection
Let's verify

```{r echo=FALSE,warning=FALSE,message=FALSE}

RedWine$alcohol_round_factor<-factor(RedWine$alcohol_round)
```

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Alchol vs. volatile acidity. Only checking the for alcohol levels
#Limiting display from 9 to 13
ggplot(aes(x=alcohol_round_factor,y=volatile.acidity),
data=subset(RedWine,(alcohol_round>8.99&alcohol_round<13.1)))+
geom_boxplot()

```

The correlation coefficient did not suggest a correlation between these
two variables and this plot does neither.
The median volatile acide goes down and than up with rising alcohol.
The spread of the volatile acide data also does not differ that much 
between alcohol levels.

## Alcohol and Density
Correlation of 0.496 was calculated for alcohol and density

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for alcohol and density, color by quality
ggplot(aes(x=alcohol,y=density),data=RedWine)+geom_jitter(alpha=1/2, position=position_jitter(h=0),aes(color=quality_factor))+
stat_smooth(method ='lm')+
#Using sprectral color palette
scale_colour_brewer(palette="Spectral", "Quality")


```

Density goes down with higher alcohol level. 

##Residual sugar and density

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for density and residual sugar, color by quality
ggplot(aes(x=residual.sugar,y=density),data=RedWine)+
scale_x_continuous(limits=c(0,8))+geom_jitter(alpha=1/2, position=position_jitter(h=0),aes(color=quality_factor))+
stat_smooth(method ='lm')+
#Using sprectral color palette
scale_colour_brewer(palette="Spectral", "Quality")

```

Density goes up with sugar.
Most of the sampples fall between 1.8 and 2.5 (approx.)



##Fixed acidity and density

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for fixed acidity and density
ggplot(aes(x=fixed.acidity,y=density),data=RedWine)+geom_jitter(alpha=1/2, position=position_jitter(h=0),aes(color=quality_factor))+
stat_smooth(method ='lm')+
#Using sprectral color palette
scale_colour_brewer(palette="Spectral", "Quality")

```

Density goes up with fixed acidity.


## Citric acid and sulphates
Looking into citric.acid (0.226) and sulphates (0.251).
These two variables are just under having a low correlation with qualtiy
I want to look at those in the linear model in the next section


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplots for quality vs. citric acid and quality vs. sulphates
plot1<-ggplot(aes(x=quality, y=citric.acid),data=RedWine)+geom_jitter(alpha=1/20)+stat_smooth(method='lm')
plot2<-ggplot(aes(x=quality, y=sulphates),data=RedWine)+geom_jitter(alpha=1/20)+stat_smooth(method='lm')
grid.arrange(plot1, plot2, ncol=2)

```

This plot shows the same weak correlation as the correleation coefficient.


### Talk about some of the relationships you observed in this part of the 
Alcohol and volatile acidity influence quality. Density and alcohol, 
Density and fixed acidity and Density and residual sugar are connected.
For these last three, I don't know if one is dependent and one is independent,
or if they just appear naturally together

### Did you observe any interesting relationships between the other features 
Alcohol and Density are related. Density is related to 
fixed acidity and residual sugar.
Apart from alcohol and volatile acidity I could not find other variables
who influence quality direclty. This surprises me, I
would have imagined that at least there can be an influence seen from density

### What was the strongest relationship you found?
Fixed acidity and density have the highest correlation coefficient.
Regarding quality, the strongest relation is still with alcohol.


# Multivariate Plots Section

The correlation suggest that alcohol and volatile acidity influence quality.

Getting the median combination of alcohol and quality in the two highest 
quality levels #and in the medium (5 and 6) levels


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Creating Medium Quality Wine Set
RedWineMQ <-(subset(RedWine, (quality==5 | quality==6)))
summary(RedWineMQ$volatile.acidity)
summary(RedWineMQ$alcohol)

```

```{r echo=FALSE,warning=FALSE,message=FALSE}

#Creating High Quality Wine set
RedWineHQ<-(subset(RedWine, quality >6))
summary(RedWineHQ$volatile.acidity)
summary(RedWineHQ$alcohol)

```

Overall numbers to compare

```{r echo=FALSE,warning=FALSE,message=FALSE}

#comparing to Wine set
summary(RedWine$volatile.acidity)
summary(RedWine$alcohol)
```

Is there a "sweet spot" for alcohol and volatile acidity?

In the graph below, the darkest dots are the wines of high quality.
The top 1% of been removed to remove outliers

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Creating a subset to exclue top 1% of acidity
RedWine99<-subset(RedWine, volatile.acidity<=quantile(RedWine$volatile.acidity, 
                                                      0.99))

#Cutting quality
RedWine99$quality.levels <- cut(RedWine99$quality, breaks = c(2.9, 5, 6, 9))
levels(RedWine99$quality.levels) = c('(2.9,5]' = 'LowQuality',
                              '(5,6]' = 'MediumQality',
                              '(6,9]' = 'HighQuality')


ggplot(aes(x=alcohol, y=volatile.acidity),data=RedWine99)+
geom_point(aes(colour=quality.levels))+
#Dark color to show high quality
scale_color_brewer(type = "div",
palette = "Purples", name="Quality")+
labs(x="Alcohol", y="Volatile Acidity") 




```


Again, this plot shows low(er) acidity and high(er) are a combination for a 
high quality wine. Put we can also see that there are points on the scatterplot
of medium quality and even low quality wines that are very close to the high 
quality wines.

The same plot, but low quality wines highlighted with darker dots

```{r echo=FALSE,warning=FALSE,message=FALSE}


ggplot(aes(x=alcohol, y=volatile.acidity),data=RedWine99)+
geom_point(aes(colour=quality.levels))+
#Dark color to show low quality
scale_color_brewer(type = "div",
palette = "Purples", name="Quality", direction = -1)+
labs(x="Alcohol", y="Volatile Acidity") 


```


High acidity seems to be a particlar sign of low quality wines.
Still there are some medium quality (and even high quality) wines
very closely located to the low quality wine.


## Linear model

Linear model will be used (u=a_1+a_2*x) for qualtiy against
alcohol, volatile acidity,density and pH.
My main interest is to find the relation between quality, alcohol and 
volatile acidity


```{r echo=FALSE}

mod1 <-lm((quality~alcohol), data=RedWine)
mod2 <-lm((quality~volatile.acidity), data=RedWine)
mod3 <-lm((quality~density), data=RedWine)
mod4 <-lm((quality~pH), data=RedWine)
mod5 <-lm((quality~citric.acid), data=RedWine)
mod6 <-lm((quality~sulphates), data=RedWine)

mtable(mod1,mod2, mod3, mod4, mod5, mod6)


```

For every 0.175 more alcohol we can expect the quality to go up by 0.017

Fir every 0.058 increase in acidity we can expect the quality to go up by 0.104
22% of variation in quality can be explained by alcohol and 15% by volatile 
acidity.

## Comparing alcohol, volatile acidity, density, sulphates and citric acid.
Boxplots to show values for high quality for:
Alcohol, Volatile Acidity, Density,citric acid and sulphates


````{r echo=FALSE,warning=FALSE,message=FALSE}
#Creating set of boxplots for High Quality Wine Segment
plot1<-ggplot(aes(y=alcohol), data=RedWineHQ)+geom_boxplot()
plot2<-ggplot(aes(y=volatile.acidity), data=RedWineHQ)+geom_boxplot()
plot3<-ggplot(aes(y=sulphates), data=RedWineHQ)+geom_boxplot()
plot4<-ggplot(aes(y=density), data=RedWineHQ)+geom_boxplot()
plot5<-ggplot(aes(y=pH), data=RedWineHQ)+geom_boxplot()
plot6<-ggplot(aes(y=citric.acid), data=RedWineHQ)+geom_boxplot()
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3)
```

Is this the ideal wine?

Loooking closer at the results for alcohol and volatile acidity.

Calculating quartile spread for alcohol and volatile acidity.


```{r}
summary(RedWineHQ$alcohol)
summary(RedWineHQ$volatile.acidity)

```

## The ideal combination?

The following graph shows all Red Wines whose Alcohol and Volatile Acidity 
levels are within the Quartils Range of High Quality Wine.
The blue dots show the wines NOT of high quality.
We can see that many lower quality wines share the acidity and alcohol 
characteristics of higher quality wine

```{r echo=FALSE,warning=FALSE,message=FALSE}

RedWineMLQ <-(subset(RedWine,(quality<=6)))

#Creating a subset for Red Wine that only contains samples with the alcohol and 
#volatile acidity of high quality wines
RWHQacid_alcohol<-(subset(RedWine,((alcohol>=10.8&alcohol<=12.2)&
(volatile.acidity>=0.3 & volatile.acidity<=0.49))))
head(RWHQacid_alcohol)


```

```{r echo=FALSE,warning=FALSE,message=FALSE}
#Alcohol vs.volatile acidty, subset of wine with high quality
#characteristcs of alcohol and volatile acidity


ggplot(aes(x=alcohol, y=volatile.acidity), data=subset(RWHQacid_alcohol,quality>5))+
geom_point(color='red', alpha=1/20)+
geom_point(aes(x=alcohol, y=volatile.acidity), 
data=subset(RWHQacid_alcohol, quality<6), color='blue')+
labs(title="Wines of lower quality showing characteristics in volatile acidity 
and alcohol of high quality wines.\nBlue dots showing low medium quality wines",
x="Quality", y="Volatile Acidity", 
color="High Quality")+
ylim(0.3, 0.49)+
xlim(10.8, 12.2)

```



Bar chart to visualize the amount of wines falling into the "ideal" acidity 
and alcohol level

```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(aes(x=quality), data=RWHQacid_alcohol)+geom_bar()
```


### Talk about some of the relationships you observed in this part of the \
High alcohol and low volatile acidity help create a high quality wine.
It seems safe to say that high acidty and low alcohol is a combination for
a low quality wine.
The opposite is not that clear. There are low and medium quality wine that 
show a combination of high alcohol and low volatile acidity.
I am assuming, that quality is measured by human testers and not by chemical
components.
There might be a component in high quality that is not captured in our data.

### Were there any interesting or surprising interactions between features?
I was suprised by the relationship of pH, fixed acidity and volatile acidity.
I would have thought that the aciditry influences pH in the same way.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
I created a linear model.
The linear model did support the result of the correlation coefficient.

------

# Final Plots and Summary


### Plot One

```{r echo=FALSE,warning=FALSE,message=FALSE}

ggplot(aes(x=quality), data=RedWine)+geom_histogram(binwidth=1, color='white')+
scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8),
labels=c("1","2","3","4","5","6","7","8"))+
labs(title="Wine samples per category")+
ylab("Quality ((score between 0 and 10)")
theme(plot.title =element_text(size=13))


```


### Description One
The plot shows the distribution of our samples in the quality levels.
Most of our samples fall into "medium" quality, 5 or 6.
The sample is rather small, so this is interesting.
Medium quality wine might not be something wine makers try to avoid.
I am assuming more consumers purchase middle quality wines. So the 
goal might be to avoid low quality wines.

### Plot Two


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Scatterplot for volatile acidity and quality. Points colored by alcohol level
ggplot(aes(y=volatile.acidity, x=quality), 
#Removing na acohol data
data=subset(RedWine, !is.na(alcohol_level) ))+
geom_jitter(alpha=1/2,position=position_jitter(h=0),
aes(color=alcohol_level))+
#Adding in line for median volatile acidity
geom_line(stat='summary', fun.y=mean)+ylim(0,1.2)+
#Changing color palette
scale_colour_brewer(palette="Set1", "Alcohol Level", direction=-1)+
scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8),
labels=c("1","2","3","4","5","6","7","8"))+
ylab(expression(Volatile~Acidity~(g/dm^{3}))) + 
xlab("Quality(score between 0 and 10)")+
labs(title="Volatile acidity vs.Quality")+
theme(plot.title =element_text(size=13))
```


### Description Two
Lower volatile acidity has a positive effect on quality.
In this plot we can also wee that higher alcohol (green and blue points) 
increase the quality. Low volatile acidity and high alcohol make a high quality
wine.

### Plot Three
```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(aes(x=alcohol, y=volatile.acidity), data=subset(RWHQacid_alcohol,quality>5))+
geom_point(color='red', alpha=1/20)+
geom_point(aes(x=alcohol, y=volatile.acidity), 
data=subset(RWHQacid_alcohol, quality<6), color='blue')+
ylab(expression(Volatile~Acidity~(g/dm^{3}))) + 
xlab("Alcohol (% by volume)")+
labs(title="Wines of lower quality showing characteristics in volatile acidity 
and alcohol of high quality wines.
Blue dots showing low and medium quality wines")+
theme(plot.title =element_text(size=13))+
ylim(0.3, 0.49)+
xlim(10.8, 12.2)

```



### Description Three
Looking only at wines of higher quality (7 and 8).
The majority of wines has a volatile acidity between 0.3 and 0.49 between and 
alcohol between 10.8 and 12.2.
Are these the sign of high quality wines? 
Maybe not, the graph above highlights in blue all low and medium wines with the 
same characteristics.
------

# Reflection

Even after investigating the data set, I feel something is missing.
I would have expected to have a clearer picture about wine quality.
Maybe this is caused by quality involving a human factor?
While it is clear that certain characteristics make a bad wine, I am 
not sure I can confidenlty say what makes a good wine.

It would have helped to include more wine samples, possibly from more regions.
Price would be a hepful data, as I am assuming in the end this is about
helping winemakers to make profit.

I mostly struggled with not knowing enough about wine. I did research about
some of the variables.
The assignment would have been easier, if I would have had domain 
expertise.

This was a good exercise in R, as I had to go back to my lesson notes and use
what I learned.
I found it challenging to create plots that summarise the data. Since starting
working on this project I am paying closer attention to plots I see in 
books, newspapers or online articles to get a better understanding what kind
of plot works for which data.
I do feel I need to learn more about using and building a model.
I was mostly relying on what I had learned in Part 1 for linear regression.


